
# Reproducible Research
## Course Project 1


### Loading and preprocessing the data



For the first part of the analysis, we need to load the data from the csv files
and load the libraries we'll need with the following:

```{r}
library(dplyr)
library(lubridate)
library(ggplot2)

df<-read.csv("activity.csv")

```


Then validate that the data is correct

```{r}
head(df)

```


After that, proceed to adjust the class of the variables and we create a 
variable for the weekday of the date:

```{r}

df$date<-ymd(df$date)

df<- df %>% mutate(weekday=weekdays(df$date))

head(df)

```


### What is mean total number of steps taken per day?



Then we proceed to group by date and sum all the steps taken by day, and the 
histogram, excluding the values "NA"

```{r}

sum.day<-df %>% group_by(date) %>%summarise(steps=sum(steps,na.rm = TRUE))

sum.day

hist(sum.day$steps, breaks = 10,xlim = c(0,25000), ylim=c(0,20), main="Sum of steps by day", xlab = "Steps by day", col="purple")

```


### What is the average daily activity pattern?



```{r}

mean.day<-round(mean(sum.day$steps),digits = 2)

median.day<-median(sum.day$steps)

```



- The mean number of steps per day is **`r mean.day`**

- The median number of steps per day is **`r median.day`**


### What is the average daily activity pattern?



For this question we group by interval and calculate the mean of every interval
and proceed to plot:

```{r}

mean.interval<-df %>% group_by(interval) %>%summarise(steps=mean(steps,na.rm = TRUE))
max.interval<-mean.interval[which.max(mean.interval$steps),]$interval
with(mean.interval,plot(interval,steps,main="Average steps by interval", type="l", col="purple" , lwd=2))



```



Then Interval where the maximum number of steps occurs is at `r max.interval`


### Imputing missing values


```{r}
na.df<-sum(is.na(df$steps))
```


The number of values NA of the data is **`r na.df`**

For this process the values of NA in the step data with the mean of steps for the
specific interval. After the substitution were made, a new histogram is made to 
show the difference with the previous one

```{r}
df2<-df
df2$steps[is.na(df2$steps)]<-mean.interval$steps[match(df2$interval[is.na(df2$steps)],mean.interval$interval)]
sum.day2<-df2 %>% group_by(date) %>%summarise(steps=sum(steps,na.rm = TRUE))
hist(sum.day2$steps, breaks = 10,xlim = c(0,25000), ylim=c(0,20), main="Sum of steps by day, NA corrected", xlab = "Steps by day", col="purple")
mean.day2<-mean(sum.day2$steps)
median.day2<-median(sum.day2$steps)


```


The mean of steps with NA values imputed is **`r (mean.day2)`** and the median is
**`r (median.day2)`** a difference of **`r round((mean.day2 - mean.day)/mean.day *100,digits=2)`%** for
the mean, and **`r round((median.day2 - median.day)/median.day *100,digits=2)`%** for median
if we introduce values for the missing parts.


### Are there differences in activity patterns between weekdays and weekends?

A new variable is added to the dataframe and a plot to evidence de differences of steps
between weekends and weekdays is shown

```{r}
df<-df %>% mutate(type.wd=ifelse(weekday=="Saturday"| weekday=="Sunday","Weekend","Weekday"))
mean.interval2<-df %>% group_by(type.wd,interval) %>%summarise(steps=mean(steps,na.rm = TRUE))
ggplot(mean.interval2,aes(interval,steps,col=type.wd))+ geom_line()+facet_wrap(~type.wd,nrow = 2,ncol = 1)+theme(legend.position = "none")

```

We can notice that the number of steps during weekends is slightly lower then on weekdays.








